---
title: "pi_curves_nls"
author: "P Harnay"
date: "2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

```{r load_packages}
#Read in required libraries
library("devtools")
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("lubridate")
library("chron")
library("plyr")
library("dplyr")
library("tidyr")
library("tidyverse")
library("broom")
library("ggpubr")
library("minpack.lm")
library("ggpmisc")

```

Modified from A. Huffmyer H. Putnam scripts  

Load metadata
```{r}
sample.info <- read.csv("Data/Spat_Info.csv", na.strings = "NA")
sample.info$Size.mm2 <- as.numeric(as.character(sample.info$Size.mm2))

#Plot Size as a function of Polyp Number
ggplot(data = sample.info, aes(x = Polyp.Number, y = Size.mm2)) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2","P")), size = 4) +
  geom_point() +
  theme_classic()

#Log transformed Plot Size as a function of Polyp Number
ggplot(data = sample.info, aes(x = log10(Polyp.Number+1), y = log10(Size.mm2+1))) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2","P")), size = 4) +
  geom_point() +
  theme_classic()
```

PLATE1
# Import data
```{r import_data}
Data <-read.csv("data/Imaging_pam/20230505_Spat_Plate1_newpoi.csv", header=TRUE, sep=",", na.string="NA") #reads in the data files
Data <- Data[, c("Date", "Time", "PAR", "ETR1","ETR2","ETR3","ETR4","ETR5","ETR6","ETR7","ETR8","ETR9","ETR10","ETR11",
      "ETR12","ETR13","ETR14","ETR15","ETR16","ETR17","ETR18",
      "ETR19","ETR20","ETR21","ETR22")]
# change format from wide to long
Data <- Data %>%  pivot_longer(., cols = starts_with("ETR"), names_to = "Spat.ID")
Data$value <- na_if(Data$value, 0)
#need to Remove all data at PAR values above where ETR hits zero
Data$PAR <- as.numeric(Data$PAR)
colnames(Data)[5] ="ETR"

```


# View Data plots
```{r}
Data %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point()+
  geom_line()+
  facet_wrap("Spat.ID")+
  theme_bw()

```

# fit all of the data together
```{r}
set.seed(123) #Random Number Generation
Data.PI <- Data %>% subset(PAR<900) # filter for PAM values from light levels <900
curve.nlsPIC <- Data.PI %>% nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) #calculate curve fit
coef(curve.nlsPIC) #examine coefficients
plot(ETR ~ PAR, data = Data.PI) #plot data
lines(0:900, predict(curve.nlsPIC, newdata = data.frame(PAR = 0:900)))
```
# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Using fixed initial values (keeping for reference): Am=0.7,  AQY=0.001, Rd=.4  

Fit each curve separately
```{r}
# filter out samples with issues in the data identified in the plots above
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR13"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR16"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR2"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR7"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR1"))

# Run many curve fits
nls_data1 <- Data.PI %>%
group_by(Spat.ID) %>%
nest(-Spat.ID) %>%
mutate(model1 = map(data, ~
nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) %>%
tidy %>%
dplyr::select(term, estimate) %>%
spread(term, estimate))) %>%
unnest(model1) %>%
mutate(Ik = Am/AQY)%>%
mutate(Date = "20230505") %>%
mutate(Plate = "Plate1") %>%
mutate(Spat.ID.PAM = paste0(Date,"_",Plate,"_",Spat.ID)) %>%
write_csv(., "Output/Plate1_pi_curve_pars_NLS_fixedparms.csv")
```

Plot curve over data points.  
```{r}
augmented <- Data.PI %>%
  filter(PAR <900) %>%
  nest(-Spat.ID) %>% 
  mutate(
    fit = map(data, ~ nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

augmented$ID <-augmented$Spat.ID
Aug <- augmented %>% separate(ID, c("Sp", "Num", "Type"))

#all colonies together
pdf("Output/Plate1_PI_Curve_Fit_Plots.pdf")
PI.curves <- Aug %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point() + 
  geom_line(aes(y=.fitted, x=PAR,group=Spat.ID))+
  theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste("ETR")))+
  theme(legend.position = "top")
PI.curves
dev.off()

```
#PLATE 2

# Import data
```{r import_data}
Data <-read.csv("Data/Imaging_pam/20230505_Spat_Plate2_newpoi.csv", header=T, sep=";", na.string="NA") #reads in the data files
Data <- Data[, c("Date", "Time", "PAR", "ETR1","ETR2","ETR3","ETR4","ETR5","ETR6","ETR7","ETR8","ETR9","ETR10","ETR11",
      "ETR12","ETR13","ETR14","ETR15","ETR16","ETR17","ETR18",
      "ETR19","ETR20","ETR21","ETR22")]
# change format from wide to long
Data <- Data %>%  pivot_longer(., cols = starts_with("ETR"), names_to = "Spat.ID")
Data$value <- na_if(Data$value, 0)
#need to Remove all data at PAR values above where ETR hits zero
Data$PAR <- as.numeric(Data$PAR)
colnames(Data)[5] ="ETR"

```


# View Data plots
```{r}
Data %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point()+
  geom_line()+
  facet_wrap("Spat.ID")+
  theme_bw()
```

```{r}
set.seed(123)
Data.PI  <- Data %>% subset(PAR<900) 
curve.nlsPIC <- Data.PI %>% nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))
coef(curve.nlsPIC)
plot(ETR ~ PAR, data = Data.PI)
lines(0:900,
      predict(curve.nlsPIC,
              newdata = data.frame(PAR = 0:900)))
```
# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Using fixed initial values  
```{r}

Data.PI <- Data.PI %>% subset(Spat.ID!="ETR12")
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR13"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR18"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR6"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR7"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR9"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR16"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))

nls_data2 <- Data.PI %>%
group_by(Spat.ID) %>%
nest(-Spat.ID) %>%
mutate(model1 = map(data, ~
nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) %>%
tidy %>%
dplyr::select(term, estimate) %>%
spread(term, estimate))) %>%
unnest(model1) %>%
mutate(Ik = Am/AQY)%>%
mutate(Date = "20230505") %>%
mutate(Plate = "Plate2") %>%
mutate(Spat.ID.PAM = paste0(Date,"_",Plate,"_",Spat.ID)) %>%
write_csv(., "Output/Plate2_pi_curve_pars_NLS_fixedparms.csv")
```

Plot curve over data points.  
```{r}
augmented <- Data.PI %>%
  filter(PAR <900) %>%
  nest(-Spat.ID) %>% 
  mutate(
    fit = map(data, ~ nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

augmented$ID <-augmented$Spat.ID
Aug <- augmented %>% separate(ID, c("Sp", "Num", "Type"))

#all colonies together
pdf("Output/Plate2_PI_Curve_Fit_Plots.pdf")
PI.curves <- Aug %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point() + 
  geom_line(aes(y=.fitted, x=PAR,group=Spat.ID))+
  theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste("ETR")))+
  theme(legend.position = "top")
PI.curves
dev.off()

```

#PLATE 3

# Import data
```{r import_data}
Data <-read.csv("Data/Imaging_pam/20230510_Spat_Plate3.csv", header=T, sep=",", na.string="NA") #reads in the data files
Data <- Data[, c("Date", "Time", "PAR", "ETR1","ETR2","ETR3","ETR4","ETR5","ETR6","ETR7","ETR8","ETR9","ETR10","ETR11",
      "ETR12","ETR13","ETR14","ETR15","ETR16","ETR17","ETR18",
      "ETR19","ETR20","ETR21","ETR22")]
# change format from wide to long
Data <- Data %>%  pivot_longer(., cols = starts_with("ETR"), names_to = "Spat.ID")
Data$value <- na_if(Data$value, 0)
#need to Remove all data at PAR values above where ETR hits zero
Data$PAR <- as.numeric(Data$PAR)
colnames(Data)[5] ="ETR"

```


# View Data plots
```{r}
Data %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point()+
  geom_line()+
  facet_wrap("Spat.ID")+
  theme_bw()
```

```{r}
set.seed(123)
Data.PI  <- Data %>% subset(PAR<900) 
curve.nlsPIC <- Data.PI %>% nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))
coef(curve.nlsPIC)
plot(ETR ~ PAR, data = Data.PI)
lines(0:900,
      predict(curve.nlsPIC,
              newdata = data.frame(PAR = 0:900)))
```
# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Using fixed initial values  
```{r}

Data.PI <- Data.PI %>% subset(Spat.ID!="ETR1")
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR14"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR17"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR4"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR6"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR9"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR18"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))

nls_data3 <- Data.PI %>%
group_by(Spat.ID) %>%
nest(-Spat.ID) %>%
mutate(model1 = map(data, ~
nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) %>%
tidy %>%
dplyr::select(term, estimate) %>%
spread(term, estimate))) %>%
unnest(model1) %>%
mutate(Ik = Am/AQY)%>%
mutate(Date = "20230510") %>%
mutate(Plate = "Plate3") %>%
mutate(Spat.ID.PAM = paste0(Date,"_",Plate,"_",Spat.ID)) %>%
write_csv(., "Output/Plate3_pi_curve_pars_NLS_fixedparms.csv")
```

Plot curve over data points.  
```{r}
augmented <- Data.PI %>%
  filter(PAR <900) %>%
  nest(-Spat.ID) %>% 
  mutate(
    fit = map(data, ~ nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

augmented$ID <-augmented$Spat.ID
Aug <- augmented %>% separate(ID, c("Sp", "Num", "Type"))

#all colonies together
pdf("Output/Plate3_PI_Curve_Fit_Plots.pdf")
PI.curves <- Aug %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point() + 
  geom_line(aes(y=.fitted, x=PAR,group=Spat.ID))+
  theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste("ETR")))+
  theme(legend.position = "top")
PI.curves
dev.off()

```


#PLATE 4

# Import data
```{r import_data}
Data <-read.csv("Data/Imaging_pam/20230510_Spat_Plate4.csv", header=T, sep=";", na.string="NA") #reads in the data files
Data <- Data[, c("Date", "Time", "PAR", "ETR1","ETR2","ETR3","ETR4","ETR5","ETR6","ETR7","ETR8","ETR9","ETR10","ETR11",
      "ETR12","ETR13","ETR14","ETR15","ETR16","ETR17","ETR18",
      "ETR19","ETR20","ETR21","ETR22")]
# change format from wide to long
Data <- Data %>%  pivot_longer(., cols = starts_with("ETR"), names_to = "Spat.ID")
Data$value <- na_if(Data$value, 0)
#need to Remove all data at PAR values above where ETR hits zero
Data$PAR <- as.numeric(Data$PAR)
colnames(Data)[5] ="ETR"

```


# View Data plots
```{r}
Data %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point()+
  geom_line()+
  facet_wrap("Spat.ID")+
  theme_bw()
```

```{r}
set.seed(123)
Data.PI  <- Data %>% subset(PAR<900) 
curve.nlsPIC <- Data.PI %>% nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))
coef(curve.nlsPIC)
plot(ETR ~ PAR, data = Data.PI)
lines(0:900,
      predict(curve.nlsPIC,
              newdata = data.frame(PAR = 0:900)))
```
# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Using fixed initial values  
```{r}

Data.PI <- Data.PI %>% subset(Spat.ID!="ETR19")
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR2"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR20"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR7"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR4"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR6"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR9"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR18"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))

nls_data4 <- Data.PI %>%
group_by(Spat.ID) %>%
nest(-Spat.ID) %>%
mutate(model1 = map(data, ~
nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) %>%
tidy %>%
dplyr::select(term, estimate) %>%
spread(term, estimate))) %>%
unnest(model1) %>%
mutate(Ik = Am/AQY)%>%
mutate(Date = "20230510") %>%
mutate(Plate = "Plate4") %>%
mutate(Spat.ID.PAM = paste0(Date,"_",Plate,"_",Spat.ID)) %>%
write_csv(., "Output/Plate4_pi_curve_pars_NLS_fixedparms.csv")
```

Plot curve over data points.  
```{r}
augmented <- Data.PI %>%
  filter(PAR <900) %>%
  nest(-Spat.ID) %>% 
  mutate(
    fit = map(data, ~ nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

augmented$ID <-augmented$Spat.ID
Aug <- augmented %>% separate(ID, c("Sp", "Num", "Type"))

#all colonies together
pdf("Output/Plate4_PI_Curve_Fit_Plots.pdf")
PI.curves <- Aug %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point() + 
  geom_line(aes(y=.fitted, x=PAR,group=Spat.ID))+
  theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste("ETR")))+
  theme(legend.position = "top")
PI.curves
dev.off()

```


#PLATE 5

# Import data
```{r import_data}
Data <-read.csv("Data/Imaging_pam/20230510_Spat_Plate5.csv", header=T, sep=";", na.string="NA") #reads in the data files
Data <- Data[, c("Date", "Time", "PAR", "ETR1","ETR2","ETR3","ETR4","ETR5","ETR6","ETR7","ETR8","ETR9","ETR10","ETR11",
      "ETR12","ETR13","ETR14","ETR15","ETR16","ETR17","ETR18",
      "ETR19","ETR20","ETR21","ETR22")]
# change format from wide to long
Data <- Data %>%  pivot_longer(., cols = starts_with("ETR"), names_to = "Spat.ID")
Data$value <- na_if(Data$value, 0)
#need to Remove all data at PAR values above where ETR hits zero
Data$PAR <- as.numeric(Data$PAR)
colnames(Data)[5] ="ETR"

```


# View Data plots
```{r}
Data %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point()+
  geom_line()+
  facet_wrap("Spat.ID")+
  theme_bw()
```

```{r}
set.seed(123)
Data.PI  <- Data %>% subset(PAR<900) 
curve.nlsPIC <- Data.PI %>% nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))
coef(curve.nlsPIC)
plot(ETR ~ PAR, data = Data.PI)
lines(0:900,
      predict(curve.nlsPIC,
              newdata = data.frame(PAR = 0:900)))
```
# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Using fixed initial values  
```{r}

Data.PI <- Data.PI %>% subset(Spat.ID!="ETR10")
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR16"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR4"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR7"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR4"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR6"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR9"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR18"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))

nls_data5 <- Data.PI %>%
group_by(Spat.ID) %>%
nest(-Spat.ID) %>%
mutate(model1 = map(data, ~
nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) %>%
tidy %>%
dplyr::select(term, estimate) %>%
spread(term, estimate))) %>%
unnest(model1) %>%
mutate(Ik = Am/AQY)%>%
mutate(Date = "20230510") %>%
mutate(Plate = "Plate5") %>%
mutate(Spat.ID.PAM = paste0(Date,"_",Plate,"_",Spat.ID)) %>%
write_csv(., "Output/Plate5_pi_curve_pars_NLS_fixedparms.csv")
```

Plot curve over data points.  
```{r}
augmented <- Data.PI %>%
  filter(PAR <900) %>%
  nest(-Spat.ID) %>% 
  mutate(
    fit = map(data, ~ nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

augmented$ID <-augmented$Spat.ID
Aug <- augmented %>% separate(ID, c("Sp", "Num", "Type"))

#all colonies together
pdf("Output/Plate5_PI_Curve_Fit_Plots.pdf")
PI.curves <- Aug %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point() + 
  geom_line(aes(y=.fitted, x=PAR,group=Spat.ID))+
  theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste("ETR")))+
  theme(legend.position = "top")
PI.curves
dev.off()

```


#PLATE 6

# Import data
```{r import_data}
Data <-read.csv("Data/Imaging_pam/20230510_Spat_Plate6.csv", header=T, sep=";", na.string="NA") #reads in the data files
Data <- Data[, c("Date", "Time", "PAR", "ETR1","ETR2","ETR3","ETR4","ETR5","ETR6","ETR7","ETR8","ETR9","ETR10","ETR11",
      "ETR12","ETR13","ETR14","ETR15","ETR16","ETR17","ETR18",
      "ETR19","ETR20","ETR21")]
# change format from wide to long
Data <- Data %>%  pivot_longer(., cols = starts_with("ETR"), names_to = "Spat.ID")
Data$value <- na_if(Data$value, 0)
#need to Remove all data at PAR values above where ETR hits zero
Data$PAR <- as.numeric(Data$PAR)
colnames(Data)[5] ="ETR"

```


# View Data plots
```{r}
Data %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point()+
  geom_line()+
  facet_wrap("Spat.ID")+
  theme_bw()
```

```{r}
set.seed(123)
Data.PI  <- Data %>% subset(PAR<900) 
curve.nlsPIC <- Data.PI %>% nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))
coef(curve.nlsPIC)
plot(ETR ~ PAR, data = Data.PI)
lines(0:900,
      predict(curve.nlsPIC,
              newdata = data.frame(PAR = 0:900)))
```
# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Using fixed initial values  
```{r}

Data.PI <- Data.PI %>% subset(Spat.ID!="ETR12")
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR18"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR21"))
Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR3"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR4"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR6"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR9"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR18"))
#Data.PI <- Data.PI %>% subset(Spat.ID!=("ETR19"))

nls_data6 <- Data.PI %>%
group_by(Spat.ID) %>%
nest(-Spat.ID) %>%
mutate(model1 = map(data, ~
nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4)) %>%
tidy %>%
dplyr::select(term, estimate) %>%
spread(term, estimate))) %>%
unnest(model1) %>%
mutate(Ik = Am/AQY)%>%
mutate(Date = "20230510") %>%
mutate(Plate = "Plate6") %>%
mutate(Spat.ID.PAM = paste0(Date,"_",Plate,"_",Spat.ID)) %>%
write_csv(., "Output/Plate6_pi_curve_pars_NLS_fixedparms.csv")
```

Plot curve over data points.  
```{r}
augmented <- Data.PI %>%
  filter(PAR <900) %>%
  nest(-Spat.ID) %>% 
  mutate(
    fit = map(data, ~ nls(ETR ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

augmented$ID <-augmented$Spat.ID
Aug <- augmented %>% separate(ID, c("Sp", "Num", "Type"))

#all colonies together
pdf("Output/Plate6_PI_Curve_Fit_Plots.pdf")
PI.curves <- Aug %>%
ggplot(aes(x=PAR, y=ETR, color=Spat.ID))+
  geom_point() + 
  geom_line(aes(y=.fitted, x=PAR,group=Spat.ID))+
  theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste("ETR")))+
  theme(legend.position = "top")
PI.curves
dev.off()

```
#viziualisation of the data point distribution 
```{r}
PI.params <- rbind(nls_data1, nls_data2, nls_data3, nls_data4, nls_data5, nls_data6)

PI.params <- PI.params[,-2]

# change format from wide to long
PI.params <- PI.params %>%  pivot_longer(., cols = Am:Ik, names_to = "metric", values_to = "value")

PI.params <- PI.params %>% subset(metric!="Rd")

PI.params$Spat.ID.PAM <- paste0(PI.params$Date, "_", PI.params$Plate, "_", PI.params$Spat.ID) 
PI.params <- left_join(PI.params, sample.info, by="Spat.ID.PAM")

PI.params <- PI.params %>%
  subset(Type!="Algae")
  
ggplot(data = PI.params, aes(x = Size.mm2, y = value)) +
  geom_point() +
  facet_wrap(~metric, scales = "free_y")
```
#some outlier are present and need to be remove from the dataset 
Remove Outliers
```{r}
Am <- PI.params %>%
  subset(metric=="Am") %>%
  subset(value<80) #%>%
 # subset(Size.mm2<6)
Am$RV <- "Electron Transport Rate"

AQY <- PI.params %>%
  subset(metric=="AQY") %>%
  subset(Size.mm2<6)
AQY$RV <- "Photochemical Efficiency"

Ik <- PI.params %>%
  subset(metric=="Ik") %>%
  subset(value<375) #%>%
  #subset(Size.mm2<6)
Ik$RV <- "Saturating Irradiance"

PI.params <- rbind(Am,AQY,Ik)

ggplot(data = PI.params, aes(x = Size.mm2, y = value)) +
  geom_point() +
  facet_wrap(~metric, scales = "free_y")

write_csv(PI.params, "Output/PI.params.csv")

ggplot(data = PI.params, aes(x = log10(Size.mm2+1), y = log10(value+1))) +
  geom_point() +
  facet_wrap(~RV, scales = "free_y")

```

#Plot metrics as a function of Spat Area
```{r}

size <- ggplot(data = PI.params, aes(x = log10(Size.mm2+1), y = log10(value+1))) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2","P")), size = 2.5) +
  facet_wrap(~RV, scales = "free_y") +
  theme_classic()
size

plots <- lapply(unique(PI.params$RV), function(param) {
  df <- subset(PI.params, RV == param)
  
  ggplot(df, aes(x = log10(Size.mm2 + 1), y = log10(value + 1))) +
    stat_poly_line(color = "cyan2", linewidth = 1, se = TRUE, 
                   fill = "darkslategray3", alpha = 0.2) +   # regression line + SE ribbon
    stat_poly_eq(use_label(c("eq", "R2", "P")), size = 4) +  # equation + R²
    geom_point(color = "darkgray", size = 2, stroke = 0.7, shape = 21, fill = "darkgray") + # clear points
    theme_classic(base_size = 14) +
    ggtitle(param)
})

plots[[1]]
plots[[2]]
plots[[3]]

ggsave("Output/electron_transport_rate.png", plot = plots[[1]], width = 8, height = 6, dpi = 300)
ggsave("Output/Photochemical_efficiency.png", plot = plots[[2]], width = 8, height = 6, dpi = 300)
ggsave("Output/Saturating_irradiance.png", plot = plots[[3]], width = 8, height = 6, dpi = 300)
```
#need to move to linear model on Saturating irradiance because only one significant after log10
#R2 explain only 4% of the data which is very low 

#Statistic of the linear model 
#Distribution of the data
```{r, warning=FALSE, message=FALSE}
#Histogram vizualisation 
hist(PI.params$Size.mm2, main = "Histogram of Respiration Rates",
     xlab = "Size.mm2", col = "lightblue", breaks = 15)

#QQ-plot
qqnorm(PI.params$Size.mm2)
qqline(PI.params$Size.mm2, col = "red", lwd = 2)
```
#The histogram and the QQ plot don't show me an very good normalisation so a shapiro test need to be run to see what it's look like 

```{r}
shapiro.test(PI.params$Size.mm2)
```
#Shapiro test has a significant value, p-value= <2.2e-16 so we can reject H0 but a transforamtion need to be operate 

```{r}
#log10 transformation 
PI.params$Size.mm2_log10 <- log10(PI.params$Size.mm2 + 1)

# Check normality after log10 transformation
shapiro.test(log10(PI.params$Size.mm2 + 1))
```

```{r}
#regression for Ik
model_Ik <- lm(log10(value) ~ Size.mm2, data = subset(PI.params, metric == "Ik"))

#Histogram vizualisation for normality of data
hist(model_Ik$residuals, main = "Histogram of Respiration Rates",
     xlab = "R.nmol.mm2.min", col = "lightblue", breaks = 15)

#QQ-plot for normality of data
qqnorm(model_Ik$residuals)
qqline(model_Ik$residuals, col = "red", lwd = 2)

#check for homoscedasticity
plot(model_Ik, which = 1)  # Residuals vs Fitted

summary(model_Ik)

```





















Plot metrics as a function of Polyp Number
```{r}
polyp <- ggplot(data = PI.params, aes(x = log10(Polyp.Number+1), y = log10(value+1))) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2","P")), size = 2.5) +
  geom_point() +
  facet_wrap(~RV, scales = "free_y")+
  theme_classic()
polyp
```

Plot photo phys as a function of surface area and size
```{r}

figure <- ggarrange(size, polyp,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2)
figure

ggsave("RAnalysis/output/photophys.jpg", figure, width = 8, height=8)

```